---
title: "homework_3"
author: "Camille Okonkwo"
date: "2024-02-09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### reading in data
```{r}
hwdata1 = read_csv("data/hwdata1.csv")
```

1. Repeat the log-rank test between the survival curves of the two
treatment groups in the myelomatosis example in this week’s lecture notes. Compute the Kaplan-Meier estimators and 95% CI of the survival functions of the two treatment groups (one Kaplan-Meier table for each group).

```{r}
library(survival)

# create dataset
myelomatosis = data.frame(
  OBS = c(1:25),
  DUR = c(8, 180, 632, 852, 52, 2240, 220, 63, 195, 76, 70, 8, 13, 1990, 1976, 18, 700, 1296, 1460, 210, 63, 1328, 1296, 365, 23),
  STATUS = c(1, 1, 1, 0, 1, 0, 1, 1, 1, 1, 1, 1, 1, 0, 0, 1, 1, 0, 0, 1, 1, 0, 1, 0, 1),
  TREAT = c(1, 0, 0, 1, 1, 0, 1, 1, 0, 0, 0, 1, 0, 0, 1, 0, 0, 1, 1, 0, 1, 1, 0, 1, 0)
)

head(myelomatosis)
```

```{r}
#log-rank test
print(survdiff(Surv(DUR, STATUS)~TREAT, data = myelomatosis),
digits=5)

library(ggsurvfit)

#plot
survfit2(Surv(DUR, STATUS)~TREAT, data = myelomatosis) |> 
  ggsurvfit() +
  add_censor_mark() +
  add_pvalue(location = "annotation", 
             caption = "Log-rank {p.value}")
```

```{r}
#K-M table

km_table = survfit(Surv(DUR, STATUS) ~ TREAT, data = myelomatosis, conf.type = "log-log")

summary(km_table)
```

2a.	Generate a graph of the survival functions in the two hormone therapy groups using `hwdata1`. Interpret the graph. (hint: What do you observe on the difference of the survival functions between the two groups?)

```{r}
library(ggsurvfit)

survfit2(Surv(rectime, censrec) ~ hormone, data = hwdata1) |> 
ggsurvfit() +
add_censor_mark() +
add_pvalue(location="annotation",
caption="Log-rank {p.value}")
```

Hormone group 1 and 2 initially have similar rates of survival, then they become more different over time. Overall, we see that those in group 2 have a higher probability of survival compared to group 1. 


2b. Test whether the survival curves for the two hormone therapy groups are the same using the log-rank test at α = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion. 
```{r}
# log-rank test
print(survdiff(Surv(rectime, censrec) ~ hormone, data = hwdata1), 
      digits = 5)
```
#### Our hypotheses:

* H0 :S1(t) = S2(t), for all t ≤ τ
* Hα :at least one of the Sk (t) is different for some t ≤ τ

#### Test statistic:

* Q_log-rank = 8.6

#### Degree of freedom: 
* df = 1

#### P-value: 
* Pr(χ21 ≥  8.6) = 0.00343 < 0.05

#### Conclusion: 

We reject H0 at the significance level 0.05. The survival curves for patients in two hormone therapy groups are significantly different.

2c. Repeat the test in Question 2b using the Wilcoxon test. Does it lead to the same conclusion? Briefly explain why the test statistic of the Wilcoxon test is smaller than that of the log-rank test in this application. 

```{r}
library(survMisc)

fit = ten(survfit(Surv(rectime, censrec) ~ hormone, data = hwdata1))
comp(fit)
knitr::kable(attributes(fit)$lrt[, c(1, 6:8)], "simple", digits = 4)
```

From the K-M survival curve,we can see that the observed values between hormone group 1 and 2 are similar at first, then become more different. Because the log rank tests weighs the difference between observed and expected number of events equally across all event times, and the Wilcoxon test weighs the difference by the number of risk, we can expect that the log-rank Q estimate will be larger than the Wilcoxon Q estimate, and the log-rank p-value will be smaller than the Wilcoxon p-value. 

2d. Categorize the tumor size into four groups using the sample quartiles (< 20, [20, 25), [25, 35), and ≥ 35). Test whether the survival functions are the same among the four groups using the generalized log-rank test at α = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion. 
```{r}
hwdata1$size_group =
  cut(hwdata1$size, 
      c(min(hwdata1$size), 20, 25, 35,
        max(hwdata1$size)),
      include.lowest = TRUE, right = FALSE)

print(survdiff(Surv(rectime, censrec) ~ size_group, data = hwdata1),
      digits = 5)
```

#### Our hypotheses:

* H0 :S1(t) = S2(t) = S3(t) = S4(t), for all t ≤ τ
* Hα :at least one of the Sk (t) is different for some t ≤ τ

#### Test statistic:

* Q_log-rank = 19.7

#### Degree of freedom: 
* df = 3

#### P-value: 
* Pr(χ23 ≥  19.7) = 2e-04  < 0.05

#### Conclusion: 

We reject H0 at the significance level 0.05. The survival curves for patients in the four groups divided by the sample quartiles of tumor size are significantly different.