---
title: "week_3_activities"
output: github_document
date: "2024-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


### Exercise 1:

##### loading in data
```{r}
whas100 = read_csv("data/whas100.csv")
```

### Part A: Take the WHAS100 dataset, create a binary variable ‘losbin’ based on length of hospital stay such that 0: los ≤ 5 and 1: los > 5.

##### creating a binary variable losbin using a binary operator
```{r}
whas100$losbin = 1*(whas100$los>5)
```

### Part B: Test whether the survival curves for the two groups defined by ‘losbin’ are the same using the log-rank test at α = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion

##### testing survival curves using log rank
```{r}
library(survival)

print(survdiff(Surv(lenfol, fstat)~losbin, data = whas100), digits = 5)
```


#### Our hypotheses:

* H0 :Slos≤5(t) = Slos>5(t), for all t ≤ τ
* Hα :Slos≤5(t) = / = Slos>5(t), for some t ≤ τ

#### Test statistic:

* Q_log-rank = 0.2262

#### Degree of freedom: 
* df = 1

#### P-value: 
* Pr(χ21 ≥ 0.2262) = 0.634 > 0.05

#### Conclusion: 

We fail to reject H0 at the significance level 0.05. The survival curves for patients stay in hospital for 5 days or less and patients there for more than 5 days are not significantly different at the 5% level.

