---
title: "week_3_activities"
output: github_document
date: "2024-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


### Exercise 1:

##### loading in data
```{r}
whas100 = read_csv("data/whas100.csv")
```

### Part A: Take the WHAS100 dataset, create a binary variable ‘losbin’ based on length of hospital stay such that 0: los ≤ 5 and 1: los > 5.

##### creating a binary variable losbin using a logical operator
```{r}
whas100$losbin = 1*(whas100$los>5)
```

### Part B: Test whether the survival curves for the two groups defined by ‘losbin’ are the same using the log-rank test at α = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion

##### testing survival curves using log rank
```{r}
library(survival)

print(survdiff(Surv(lenfol, fstat)~losbin, data = whas100), digits = 5)
```


#### Our hypotheses:

* H0 :Slos≤5(t) = Slos>5(t), for all t ≤ τ
* Hα :Slos≤5(t) = / = Slos>5(t), for some t ≤ τ

#### Test statistic:

* Q_log-rank = 0.2262

#### Degree of freedom: 
* df = 1

#### P-value: 
* Pr(χ21 ≥ 0.2262) = 0.634 > 0.05

#### Conclusion: 

We fail to reject H0 at the significance level 0.05. The survival curves for patients stay in hospital for 5 days or less and patients there for more than 5 days are not significantly different at the 5% level.


### Exercise 2:


### Part A: Create a categorical variable ‘bmicat’ using the sample quartiles of bmi (< 23.54, [23.54, 27.19), [27.19, 30.35), and ≥ 30.35)

```{r}
whas100$bmicat = cut(whas100$bmi,
c(min(whas100$bmi),23.54,27.19,30.35,
max(whas100$bmi)),
include.lowest=T,right=F)
```

### Part B: Test whether the survival functions are the same among the four groups using the generalized log-rank test at α = 0.05. Give the null and alternative hypothesis, test statistic, degrees of freedom, p-value, and conclusion.

```{r}
print(survdiff(Surv(lenfol, fstat)~bmicat, data=whas100),
digits=5)
```

#### Our hypotheses:

* H0 :S1(t) = S2(t) = S3(t) = S4(t), for all t ≤ τ
* Hα :at least one of the Sk (t) is different for some t ≤ τ

#### Test statistic:

* Q_log-rank = 8.633

#### Degree of freedom: 
* df = 3

#### P-value: 
* Pr(χ23 ≥ 8.633) = 0.0346 < 0.05

#### Conclusion: 

We reject H0 at the significance level 0.05. The survival curves for patients in the four groups divided by the sample quartiles of bmi are significantly different.